Reconstructing the last known movements of one of Nature's giants

### 2017-03-30 manuscript/blue-whale-ms.tex
Edits to abstract

@@

\mstitlepage

\parindent = 1.5em
\addtolength{\parskip}{.3em}
% Abstract 200 -300 words max
\section{Abstract}

The blue whale, *Balaenoptera musculus*, is the largest animal to have ever lived, and one of the most heavily influenced by human actions. 
Over 50 years since the International Whaling Commission banned hunting of blue whales, North Atlantic stocks are  still only at \% of pre-whaling levels.
To conserve this endangered species we need to understand its distribution, behaviour and ecology, however, high resolution data on whale ecology are scarce, particularly for the North Atlantic population. 
Here we use baleen samples from the Natural History Museum's iconic blue whale, along with new movement models, to reconstruct the movements of the whale over the last three years of its life.
We show that the whale migrated rather than remaining resident in the northern Atlantic year round. 
This is the first clear quantitative evidence of historical migrations in the species

Museum collections provide a unique opportunity to infer animal movements with non invasive methods.

@@

\section{Methods (condensed)}
%<3000 words

@@

@@

\section{Acknowledgements}
This work was funded by the British Ecological Society (grant: 5771/6815). 
We thank XXX for comments on drafts.

\section{Author contributions}

\section{Author information}

@@

\bibliographystyle{mee2}
\bibliography{darkside}

>>>>

%Contact: h.gee#nature.com*

@@

%\mstitlepage

\parindent = 1.5em
\addtolength{\parskip}{.3em}
% Abstract 200 -300 words max
\section{Abstract}
% general topic intro - 1 sentence
% intro to field - 2 sentences
% what we find/show - 1 sentence
% implications - 2-3 sentences. Results in context, how has the paper moved the field forward

Over 50 years since the International Whaling Commission banned the hunting of blue whales (\textit{Balaenoptera musculus}), North Atlantic stocks are still at only X\% of pre-whaling levels.
Conservation efforts are hampered by the scarcity of detailed data on blue whale distribution, behaviour and ecology, for both historical and present-day populations.
Here we use $\delta^{13}C$ stable isotopes from the baleen of the Natural History Museum's iconic blue whale, along with new movement models, to extract high resolution movement behaviour data for an animal that died over 125 years ago. 
We reconstructed the movements of the blue whale over the final years of its life, and show that it migrated rather than remaining resident in the northern Atlantic year round. 
This is the first quantitative evidence of historical migrations in blue whales.
Our new model provides a way to investigate migration across marine vertbrates and highlights the unique opportunities provided by museum collections for inferring past animal movements.
% About 50 words more

@@

\section{Methods (condensed)}
%<3000 words
% no figures no tables

@@

%\bibliographystyle{nature}
%\bibliography{whale}

Supplementary Information is linked to the online version of the paper at www.nature.com/nature.

@@

\section{Acknowledgements}
% adjectives not allowed!
This work was funded by the British Ecological Society (grant: 5771/6815). 
We thank XXX for comments on drafts.

\section{Author contributions}
% rough draft at present feel free to edit
CT, NC, AJ, RS, and K collected baleen from the NHM collections. 
K and CT extracted and analysed samples.
CT created the model.
EC and NC helped parameterise the model.
AJ created the figures. 
NC wrote the paper.
All authors edited drafts and approved the final version.

\section{Author information}
Reprints and permissions information is available at www.nature.com/reprints.
The authors declare no competing financial interests.
Correspondence and requests for materials should be addressed to natalie.cooper@nhm.ac.uk.
% One person only allowed

@@
